---
title: "Note Frequencies in Viper Songs"
author: "The Viper"
date: "April 30, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)

```

## Intro

Musical improvisation and composition can seem like a mysterious thing -- and quite a leap from simply playing pre-scripted notes on a page of written music. Where does it come from?

And, yet, most of us do it pretty un-self-consicously all the time as we hum, whistle, or la-de-da sing to ourselves -- or, in my case, to my dog and cat when I'm feeding them in the morning -- and we're never at a loss for what note to play next. How is that we seem know without thinking about it much what things are going to sound right as we spontaneously compose like this? And how far is it from being able to whistle, hum, or sing to being able to translate that act of improvisatory composition to an instrument?

This project tries to take one small step for a Viper -- but one giant leap for Viper-kind -- toward demystifying the process of how to make stuff up. I hope to show that at some really basic level, you're not actually working without a net out there on the highwire once you step off the platform of written notes. In fact, if you know even a little bit about the key, the chords, and the style of the music you're playing, you'll find a number of more-or-less slip-proof platfroms serving as landings all along the rope.

So let's take that first step!

## Notes and Notes and Notes and Notes

Here's the first couple of lines from a new song that....

ADD PNG OF MUSIC

Viper composed in G (how do we know). Key signature.

But your ear and your brain can tell you -- If I sang and asked you what, you'd get right (Levitin)

Here's we'll let some simple machine learning do the work of your ears and your brain.

I've gone through and counted how many times each note appears and recorded the results in a spreadsheet at <https://github.com/ryanjerving/note-frequencies/blob/master/datasets/another-day-frequency.csv>:

```{r another-day}

ad <- read.csv("https://raw.githubusercontent.com/ryanjerving/note-frequencies/master/datasets/another-day-frequency.csv")

ggplot(ad, aes(x = Note, y = Frequency, fill = Frequency)) + geom_col()


```

Analysis -- what does this tell you about your choices: which notes are landing notes, which notes are tightrope notes.

## The Question of Style

Winnebago Bay

```{r winnebago-bay, echo=FALSE}


```

## Then There's This

The Monsters are Coming
